Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2020-12-24 spectral radius of matrix new method/thesis/article121420.tex
122420 adoption

Below is our first result, in which the first condition implies the first $n-1$ columns of $C$ major to
 the columns of $C'$, and the $(k,n)$-sum column of $C$ is also major to $C'$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative eigenvectors which are $K$-rooted. And the forth condition is simpler
  but with the same meaning with Theorem~\ref{pre_thm}
\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$
\end{enumerate} %\qed
\end{thm}
%

@@

    Here we can summarize the facts we know so far as the following:
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0,$
    \end{enumerate}
    which come from Theorem~\ref{thm_main}.

    By Lemma~\ref{l_diag}, for certain $d$, if $C'+dI$ is $K$-rooted, then it has
    a $K$-rooted eigenvector with its spectral radius $\lambda + d$. $C'$ would share the same
    eigenvector with $C'+dI$ and has eigenvalue $\lambda$. So $C'+dI$ and $C+dI$ meet the
    conditions of Theorem~\ref{thm_main}, and we can show that $\rho(C' + dI) \geq \rho(C +dI)$ and
    then $\rho(C') \geq \rho(C)$  %\qed
\end{proof}

\subsection{Example}
    For the following $4\times 4$ matrix
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    and corresponding directed graph, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_C.PNG}
    \end{center}
    we choose the matrix
    $$C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 & 0\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    which satisfies the assumptions in Theorem~\ref{thm:conclusion}. Hence $\rho(C)\leq\rho(C')$. Actually, both of the values of $\rho(C)$ and $\rho(C')$ are close to 2.511547, and we have $\rho(C')-\rho(C)>0$ by calculating on computer.
    Moreover, the corresponding directed graph of matrix $C'$ is, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_Cprime.PNG}
    \end{center}

>>>>

Below is our first result, in which the first condition implies the first $n-1$ columns of $C$ major to
 the columns of $C'$, and the $(k,n)$-sum column of $C$ is also major to $C'$. The second and the third condition
 suggest that $C$ and $C'$ have nonnegative eigenvectors which are $K$-rooted. And the forth condition is simpler
  but with the same meaning with Theorem~\ref{pre_thm}
\begin{thm}\label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii} there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar
 $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii}there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar
 $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv}$v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
    \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
    \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $ with $v_i\ne 0 $;
    \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$
\end{enumerate} %\qed
\end{thm}
%

@@

    Here we can summarize the facts we know in theorem \ref{thm_main} so far as the following:
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$
         such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0,$
    \end{enumerate}
    which come from Theorem~\ref{thm_main}.

    By Lemma~\ref{l_diag}, for certain $d$, if $C'+dI$ is $K$-rooted, then it has
    a $K$-rooted eigenvector with its spectral radius $\lambda + d$. $C'$ would share the same
    eigenvector with $C'+dI$ and has eigenvalue $\lambda$. So $C'+dI$ and $C+dI$ meet the
    conditions of Theorem~\ref{thm_main}, and we can show that $\rho(C' + dI) \geq \rho(C +dI)$ and
    then $\rho(C') \geq \rho(C)$

    As the same reason, we have $\lambda=\lambda'$ such case
    if and only if
    \begin{enumerate}[label=(\alph*)]
        \item \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$
         and $v'_n\not=0;$
        \item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $
         with $v_i\ne 0 $;
        \item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$
    \end{enumerate} %\qed

\end{proof}

\subsection{Example}
    For the following $4\times 4$ matrix
    $$C=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 0 & 1\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    and corresponding directed graph, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_C.PNG}
    \end{center}
    we choose the matrix and fix $k = 3$
    $$C'=\begin{pmatrix}
    0 & 0 & 1 & 1\\
    1 & 0 & 1 & 0\\
    1 & 1 & 0 & 1\\
    1 & 1 & 1 & 0
    \end{pmatrix},$$
    Based on condition of theorem \ref{thm:conclusion},
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ 
        such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0.$
    \end{enumerate}
    Here $n=4,k=3$ so rewrite the condition of theorem \ref{thm_main} as following:
    \begin{enumerate}[label=(\roman*)]
        \item  $C[[4]|[3]]\leq C'[[4]|[3]]$ and $c_{i3}+c_{i4}\leq c'_{i3}+c'_{i4}$ for all $1\leq i\leq 4$;
        \item  there exists a $K$-rooted vector $v'=(v'_1, v'_2, v'_3, v'_4)^T$ and a scalar $\lambda'\in \mathbb{R}$
        such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
        \item there exists a nonnegative vector $v^T=(v_1, v_2, v_3, v_4)$ and a scalar $\lambda\in \mathbb{R}$ 
        such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
        \item $v^Tv'>0.$
    \end{enumerate}
    Then we check these conditions one by one,
    \begin{enumerate}[label=(\roman*)]
    \item 
    $$C[[4]|[3]]=\begin{pmatrix}
        0 & 0 & 1\\
        1 & 0 & 0\\
        1 & 1 & 0\\
        1 & 1 & 1
        \end{pmatrix},$$
    $$C'[[4]|[3]]=\begin{pmatrix}
        0 & 0 & 1\\
        1 & 0 & 1\\
        1 & 1 & 0\\
        1 & 1 & 1
        \end{pmatrix},$$
    thus,$C[[4]|[3]]\leq C'[[4]|[3]].$
    Write $c_{i3}+c_{i4}$ and $c'_{i3}+c'_{i4}$, for $1\leq i\leq 4$, as two row vectors $v_c$ and $v_{c'}$.
    $v_c = (s_1, s_2, s_3, s_4), s_i = c_{i3}+c_{i4},$ so $v_c = (2,1,1,1)$
    $v'_c = (s'_1, s'_2, s'_3, s'_4), s'_i = c'_{i3}+c'_{i4},$ so $v'_c = (2,1,1,1),$
    so $c_{i3}+c_{i4} \leq c'_{i3}+c'_{i4}.$
    \item 
    Let $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$.
    Then $\lambda' = 2.511547141694532$ and associated eigenvector $v'=(1, 0.8981609516297208, 1.255773570847266, 1.255773570847266)^T$.
    \item 
    Let  $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$.
    Then $\lambda = 2.511547141694532$ and associated left eigenvector $v^T=(1, 0.715225238435091, 1, 0.7963219032594416)$.
    \item 
    $v^T * v' = (1, 0.715225238435091, 1, 0.7963219032594416) \\
    * (1, 0.8981609516297208, 1.255773570847266, 1.255773570847266)^T
    \\ = 3.89816095162972 >0.$

    \end{enumerate}

    These facts satisfy the assumptions in Theorem~\ref{thm:conclusion}. Hence $\rho(C)\leq\rho(C')$. Actually, both of
     the values of $\rho(C)$ and $\rho(C')$ are close to 2.511547, and we have $\rho(C')-\rho(C)>0$ by calculating on
     computer.\cite[sage]{sage}
    Moreover, the corresponding directed graph of matrix $C'$ is, \cite[sage]{sage}
    \begin{center}
    \includegraphics{graph_Cprime.PNG}
    \end{center}
